---
title: "Project proposal: Site Fidelity and Inter-pool Movement of Spotted Salamanders by the Sea"
author: "Georgia Lattig"
date: "02/14/24"
output: github_document
---

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)
library(readr)
```

## 1. Introduction

The aim of this project is to organize, synthesize and begin to analyze data from the past nine seasons of data collection for the Otter Point Salamander project. The Otter Point Salamander project (unofficial title) is a long-term monitoring project that is focused on a population of Spotted Salamanders (Ambystoma maculatum) that breeds on a rocky headland by the Atlantic Ocean. This population is unique because they breed in a series of freshwater pools in the granite headland, so close to the ocean that the pools are occasionally flushed with full-strength seawater during stochastic storm events. The project was initiated by Dr. Stephen Ressel in 2015 with efforts to quantify the environmental salinities that these salamanders were experiencing during the breeding season. In 2017, Dr. Ressel implemented a mark-recapture study to gauge the rate of salamander returns to this unusual breeding site. 

In 2022, a new and exciting aspect was added to the study by COA alum Jasper White. The original methods for mark-recapture were through the use of Visible Implant Elastomer (VIE) tags. To tag salamanders with VIE tags, we would inject a small amount of colorful, fluorescing polymer under the salamander's skin at the base of their tail. Unique colors were used for each unique year so, for example, if a salamander had a yellow tag that means it was initially marked in 2017. There are limits to this type of tagging, however, because we were only able to mark cohorts from a given year rather than uniquely identify individual salamanders. In 2022, Jasper implemented PIT-tagging as a method of marking individual salamanders. PIT tags are small microchip tags that are inserted into the body cavity of the salamanders, assigning each individual a unique barcode of numbers. PIT-tagging continued last spring (2023) and will continue this coming spring (2024). 

Both the quantity and type of data that has been collected on the Otter Point population for the past nine years opens many possible questions. With my final project, I am specifically interested in beginning to look at the site fidelity of individuals across seasons and the movement of individuals between breeding pools within each season. For both of these questions, I will be working with data from more recent years (2022-2023) and will hopefully set up the code so that this coming spring's data can easily be incorporated. Another major goal of mine in working with these data is that, in the lengthy process of data-cleaning and tidying that this project will require, I will come up with a standardized method of data collection and ideally design a data sheet for use in the field for future seasons.


I will work with one dataset that contains 9 years (currently separate data sheets) of data. There is a lot of data cleaning that still needs to be done because there are 10 different sheets, one sheet for each season, except for the 2022 season which for some reason has two. Each sheet has a slightly different number and configuration of variables. The variables are as follows:

#### op15
`Date`, `Pool Number`, `Salinity(ppt)`, `Second Salinity Noted (lower)`, `Temperature (C)`, `Total`, `Unmarked`, `Red (2021/2020)`, `Orange (2019)`, `Pink (2018)`, `Yellow (2017)`, `Dead Individuals`, `Eggmasses Present`, `Notes`

#### op16
`Date`, `Pool Number`, `Salinity(ppt)`, `Second Salinity Noted (lower)`, `Temperature (C)`, `Total`, `Unmarked`, `Red (2021/2020)`, `Orange (2019)`, `Pink (2018)`, `Yellow (2017)`, `Dead Individuals`, `Eggmasses Present`

#### op17
`Date`, `Pool Number`, `Salinity(ppt)`, `Second Salinity Noted (lower)`, `Temperature (C)`, `Total`, `Unmarked`, `Red (2021/2020)`, `Orange (2019)`, `Pink (2018)`, `Yellow (2017)`, `Dead Individuals`, `Eggmasses Present`

#### op18
`Date`, `Pool Number`, `Salinity(ppt)`, `Second Salinity Noted (lower)`, `Temperature (C)`, `Total`, `Unmarked`, `Red (2021/2020)`, `Orange (2019)`, `Pink (2018)`, `Yellow (2017)`, `Dead Individuals`, `Eggmasses Present`, `Returns Per Night`, `Color`, `Top # of Returns`

#### op19
`Date`, `Pool Number`, `Salinity(ppt)`, `Second Salinity Noted (lower)`, `Temperature (C)`, `Total`, `Unmarked`, `Red (2021/2020)`, `Orange (2019)`, `Pink (2018)`, `Yellow (2017)`, `Dead Individuals`, `Eggmasses Present`, `Returns Per Night`, `Color`

#### op20
`Date`, `Pool Number`, `Salinity(ppt)`, `Second Salinity Noted (lower)`, `Temperature (C)`, `Total`, `Unmarked`, `Red (2021/2020)`, `Orange (2019)`, `Pink (2018)`, `Yellow (2017)`, `Dead Individuals`, `Eggmasses Present`, `extra..14`, `Returns Per Night`, `Color`

#### op21
`Date`, `Pool Number`, `Salinity(ppt)`, `Second Salinity Noted (lower)`, `Temperature (C)`, `Total`, `Unmarked`, `Red (2021/2020)`, `Orange (2019)`, `Pink (2018)`, `Yellow (2017)`, `Dead Individuals`, `Eggmasses Present`, `Returns Per Night`, `Color/Year`, `Top # of Returns`

#### op22
`date`, `recapture`, `pool_num`, `pit_num`, `image`, `color`, `sex`, `year`, `air_temp_c`, `salinity_ppt`, `pool_temp_c`, `notes`, `extra..13`, `extra..14`

#### sals22
`date`, `svl`, `pool_num`, `pit_num`, `image`, `sex`, `color`, `year`, `air_temp_c`, `salinity_ppt`, `pool_temp_c`, `spermatophores`, `eggs`, `notes`

#### op23 
`date`, `recapture`, `pool_num`, `new_pit`, `pit_num`, `color`, `sex`, `year`, `air_temp_c`, `salinity_ppt`, `pool_temp_c`, `2nd_pool_temp_c`, `notes`


## 2. Data

```{r load-data, message = FALSE}
op15 <- read_csv("/cloud/project/data/otter point - 2015.csv")
op16 <- read_csv("/cloud/project/data/otter point - 2016.csv")
op17 <- read_csv("/cloud/project/data/otter point - 2017.csv")
op18 <- read_csv("/cloud/project/data/otter point - 2018.csv")
op19 <- read_csv("/cloud/project/data/otter point - 2019.csv")
op20 <- read_csv("/cloud/project/data/otter point - 2020.csv")
op21 <- read_csv("/cloud/project/data/otter point - 2021.csv")
op22 <- read_csv("/cloud/project/data/otter point - 2022.csv")
sals22 <- read_csv("/cloud/project/data/otter point - 2022_sals.csv")
op23 <- read_csv("/cloud/project/data/otter point - 2023.csv")
```

#### Data Skim

To give you an idea of the data in its current form, I glimpsed the first data sheet from 2015 and the most recent data sheet from last season (2023).

```{r skim-data}
glimpse(op15)
glimpse(op23)
```

## 3. Ethics review

Data collection for this project is done in partnership with Acadia National Park and has gone through the necessary review and appeals (IACUC). The permit number is _________. Brief statement about data collection methods, how it affects the salamanders, and the implications of the research : WHY it is being done if it is stressful to the wildlife.

## 4. Data analysis plan

Below are a few of the questions I am most interested in answering through this project:

* Which pools are the most frequently visited across all seasons? (confirm Steve's assumption)
* Rate of salamander returns using VIE color tags for year cohorts (n = 9 seasons): What is the proportion of returns to new recruits each season?
* Site/Pool Fidelity of individual salamanders across 2 seasons (2022, 2023): are individual salamanders returning to the same pools each season?
* Inter-pool movement of individual salamanders across 2 seasons (2022, 2023): how much are individual salamanders moving between pools within a given season?
* What is the most efficient way to collect these data in the field?

