---
title: "Salamanders by the Sea: Site Fidelity of Spotted Salamanders Breeding in Rocky Cliff Pools by the Atlantic Ocean"
author: "Georgia Lattig"
date: "03/01/24"
output: github_document
---
```{r install-packages, include = FALSE}
# install.packages("tidyverse")
# install.packages("broom")
# install.packages("readr")
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("lubridate")
# install.packages("igraph")
# install.packages("ggnetwork")
# install.packages("intergraph")
# install.packages("glmnet")
# install.packages("parsnip")
# install.packages("tidymodels")
# install.packages("ggridges")
# install.packages("hrbrthemes")
# install.packages("viridis")
# install.packages("plotly")
# install.packages("heatmaply")
```

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(igraph)
library(ggnetwork)
library(intergraph)
library(glmnet)
library(parsnip)
library(tidymodels)
library(ggridges)
library(hrbrthemes)
library(viridis)
library(plotly)
library(heatmaply)
```

```{r load-data, include = TRUE}
pools <- read_csv("/Users/georgialattig/salamanders by the sea/data/tidy_data/pools.csv")
sals <- read_csv("/Users/georgialattig/salamanders by the sea/data/tidy_data/sals.csv")
```
## Data Processing

```{r assigning-first_year-22-23}
(p22 <- sals %>% 
    filter(pit_year == "2022"))

sals <- sals %>% 
  mutate(first_year = case_when(
    pit %in% pits22 ~ "2022",
    .default = first_year
  ))

# Separate out the PIT tags that only appeared in 2023 by using an anti_join with individuals22 

pits23 <- sals %>%
  mutate(pit = as.character(pit)) %>%
  anti_join(individuals22, by = "pit")

(p23 <- sals %>% 
  filter(pit %in% pits23$pit))

sals <- sals %>% 
  mutate(first_year = case_when(
    pit %in% p23 ~ "2023",
    .default = first_year
  ))

sals <- sals %>% 
  mutate(color = case_when(
    first_year == "2017" ~ "yellow",
    first_year == "2018" ~ "pink",
    first_year == "2019" ~ "orange",
    first_year == "2020_2021" ~ "red",
    first_year == "2022" ~ "blue",
    first_year == "2023" ~ "green"
  ))
```

To date there have been 249 unique PIT tags deployed within the Otter Point population of salamanders. Below I ran some code to see which PIT tags, if any, had multiple years/colors associated with them. Out of 249 PIT tags, 46 are associated with more than one year (first year observed/marked)/color. 

```{r check-pits}
sals %>% 
  group_by(first_year) %>% 
  count()

sals %>% 
  group_by(pit) %>% 
  count(first_year)

#This code revealed that there is an erroneous pit tag 1299 (I think this is actually the IMG file number). Keep this in mind as I'll have to omit the observation from my visualizations/analysis.

sals %>% 
  count(pit)
```

Though this is a major pain, I am going to look through the data for each of these 46 PIT tags to determine which color/year should be associated with them. Then I will create a massive case_when function to recode the PIT tags as necessary. I will recode based on information about year of first PIT tag (these are concrete data) or based on the most frequent color tag recorded across seasons. In cases with equal numbers of observations of different color tags, I looked for additional information in the notes column or else trusted the 2022 observation because of increased sampling effort in 2022. Reasons are given for each decision within the code below.

```{r recode-pit-year}
sals <- sals %>% 
  mutate(first_year = case_when(
    pit == "3216" ~ "2020_2021", #once recorded as pink in 2023 but all other times, across 2 seasons, red
    pit == "3218" ~ "2022", #recorded as new in 2023 but was first tagged with PIT in 2022; was NOT double-tagged
    pit == "3229" ~ "2020_2021", #once recorded as pink in 2023 but all other times, across 2 seasons, red
    pit == "3234" ~ "2022", #first tagged with PIT in 2022
    pit == "3237" ~ "2017", #yellow tag likely missed in first/only encounter in 2022, all other observations are yellow
    pit == "3244" ~ "2018", #recorded red in 2022 but accompanying notes read "kind of pink VIE" + recorded pink in 2023
    pit == "3248" ~ "2020_2021", #once recorded as pink in 2023 but all other times, across 2 seasons, red
    pit == "3262" ~ "2018", #once recorded as orange in 2023 but all other times, across nights, pink
    pit == "3273" ~ "2020_2021", #no color recorded in 2022 but multiple nights in 2023 recorded red
    pit == "3275" ~ "2022", #no color recorded when first PIT tagged 2022 nor all other times except one "yellow" in 2023 ***
    pit == "3284" ~ "2022", #first tagged with PIT in 2022
    pit == "3289" ~ "2018", #recorded pink 4 nights in 2022 and red 3 nights in 2023...
    pit == "3304" ~ "2019", #recorded 2x orange in 2022 including accompanying notes about the orange tag, 2023 "might not be pink, hard to tell"
    pit == "3316" ~ "2020_2021", #1x red in 2022, 1x orange in 2023; trusted 2022 observation due to larger field crew and "effort"
    pit == "3321" ~ "2022", #no color recorded when first PIT tagged 2022 nor other times except one "yellow" in 2023 ***
    pit == "3322" ~ "2020_2021", #once recorded as orange in 2023 but all other times, across nights, red
    pit == "3328" ~ "2020_2021", #once no color recorded in 2023 but all other times, across nights, red
    pit == "3339" ~ "2017", #once recorded red in 2023 but all other times, across nights, yellow
    pit == "3343" ~ "2022", #first tagged with PIT in 2022
    pit == "3345" ~ "2022", #first tagged with PIT in 2022, all observations no color except one orange record in 2023
    pit == "3358" ~ "2022", #first tagged with PIT in 2022
    pit == "3359" ~ "2020_2021", #once recorded orange in 2023 but all other times, across nights, red
    pit == "3360" ~ "2018", #no color recorded but notes say pink
    pit == "3361" ~ "2019", #no color recorded when first PIT tagged in 2022 nor other 2022 observations but twice recorded as orange in 2023
    pit == "3363" ~ "2020_2021", #once no color recorded in 2023 but all other times, across seasons, red
    pit == "3370" ~ "2019", #recorded orange 4x in 2022, once pink and twice red in 2023
    pit == "3374" ~ "2017", #yellow tag likely missed in first/only encounter in 2022, all other observations are yellow
    pit == "3375" ~ "2017", #recorded yellow across 3 nights in 2023, no colors recorded in all prior observations ***
    pit == "3377" ~ "2022", #UV light forgotten so no color tag but first tagged with PIT 2022
    pit == "3380" ~ "2022", #first tagged with PIT in 2022
    pit == "3382" ~ "2022", #first tagged with PIT in 2022
    pit == "3383" ~ "2022", #first tagged with PIT in 2022
    pit == "3393" ~ "2022", #first tagged with PIT in 2022
    pit == "3396" ~ "2020_2021", #once recorded pink in 2023 but all other times, across seasons, red
    pit == "3399" ~ "2022", #no color recorded when first PIT tagged in 2022 nor all other 2022 observations except 2x recorded as red in 2023 with accompanying notes "may not be red"
    pit == "3401" ~ "2020_2021", #no color recorded when first PIT tagged in 2022 and only once recorded "red" but additional notes from another night say "has color tag but couldn't read it" ***
    pit == "3402" ~ "2019", #once recorded pink in 2023 but all other times, across nights, orange
    pit == "3403" ~ "2020_2021", #once recorded pink in 2023 but all other times, across seasons, red
    pit == "4741" ~ "2018", #always recorded pink
    pit == "4752" ~ "2019", #most often recorded orange
    pit == "4766" ~ "2023", #first tagged with PIT in 2023
    pit == "4797" ~ "2020_2021", #possible that the red color tag was missed when first PIT tagged 2023
    pit == "4832" ~ "2020_2021", #possible that the red color tag was missed when first PIT tagged 2023, all other times red
    pit == "4838" ~ "2023", #first tagged with PIT in 2023
    pit == "4842" ~ "2023", #first tagged with PIT in 2023
    pit == "4854" ~ "2023", #first tagged with PIT in 2023
    pit == "4880" ~ "2023", #first tagged with PIT in 2023
    pit == "4883" ~ "2023", #first tagged with PIT in 2023
    pit == "4888" ~ "2023", #first tagged with PIT in 2023
    pit == "4898" ~ "2020_2021", #possible that the red color tag was missed when first PIT tagged 2023, all other times red
    .default = first_year
  ))

sals <- sals %>% 
  mutate(color = case_when(
    first_year == "2017" ~ "yellow",
    first_year == "2018" ~ "pink",
    first_year == "2019" ~ "orange",
    first_year == "2020_2021" ~ "red",
    first_year == "2022" ~ "blue",
    first_year == "2023" ~ "green"
  ))
```


```{r recode-pit-year-variable}
sals <- sals %>% 
  mutate(pit = as.character(pit)) %>% 
  mutate(pit_year = as.character(pit_year)) %>% 
  mutate(pit_year = case_when(
    pit %in% p23$pit ~ "2023",
    pit %in% p22$pit ~ "2022",
    .default = pit_year
  ))
```


```{r check-first-year-again, include = FALSE}
sals %>% 
  group_by(pit) %>% 
  count(first_year)

#Below I removed the erroneous observation of PIT #1299 from the dataset (a single observation on 03/31/22)
sals <- sals %>% 
  filter(pit != "1299" |
        is.na(pit)) %>% 
  mutate(pit = as.character(pit)) %>% 
  mutate(pool = as.character(pool))
```

```{r create-observation-year-variable}
sals <- sals %>% 
  mutate(date = as.Date(date)) %>% 
  mutate(year = as.numeric(format(date,'%Y'))) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(year = case_when(
    year == "23" ~ "2023",
    .default = year))
```

```{r pit-counts-by-year}
sals %>% 
  group_by(year) %>% 
  count(pit)

sals %>% 
  count(pit)
#The code above reveals that there are 249 uniquely PIT tagged individuals

sals %>% 
  group_by(pit) %>% 
  count()

#The code below creates proportions of visits at n/sum(n) breeding pools for each individual 
sals %>% 
  group_by(pit) %>% 
  count(pool) %>% 
  mutate(pool_prop = n/sum(n))
```
```{r visualize-most-popular-breeding-pools}
sals %>% 
  drop_na(pit) %>% #Only look at tagged individuals
  group_by(pit) %>% #Will grouping by PIT remove duplicate observations?
  group_by(year) %>% 
  count(pool) %>% 
  mutate(pool = factor(pool, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16"))) %>%
  ggplot(aes(pool, n), fill = first_year) + #Okay what am I actually plotting here?
  geom_col() + 
  facet_wrap(~ year) +
  labs(title = "Breeding Pools Used by PIT Tagged Salamanders",
       subtitle = "Across Two Seasons",
       x = "Pool",
       y = "PIT Tagged Individuals Observed")

sals %>% 
  group_by(year) %>% 
  group_by(pit) %>% 
  count(pool)

pools <- pools %>% 
  mutate(date = as.Date(date)) %>% 
  mutate(year = as.numeric(format(date,'%Y'))) 

pools <- pools %>% 
  mutate(year = as.character(year)) %>% 
  mutate(year = case_when(
    year == "20" ~ "2020",
    year == "21" ~ "2021",
    .default = year)) 

pools %>% 
  filter(year %in% c("2016","2017","2018","2019","2020","2021")) %>% 
  mutate(pool = factor(pool, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16"))) %>%
  drop_na(pool) %>% 
  ggplot(aes(pool, total), fill = first_year) + #Okay what am I actually plotting here?
  geom_col() + 
  facet_wrap(~ year) +
  labs(title = "Breeding Pools Used by Salamanders",
       subtitle = "Across Seven Seasons",
       x = "Pool",
       y = "Total Salamander Observations")
##Change the units here as they are adding up "totals"... Does that work?

pools %>% 
  group_by(pool) %>% 
  count() %>% 
  arrange(n)
```

```{r observations-per-year}
sals %>% 
  group_by(year) %>% 
  count()
#There were more salamander observations than just 291 but the data in 2023 did not record untagged/uncaptured salamanders as their own rows unlike 2022 data (something to fix going forward?)
```
## Modeling Site Fidelity

### Data Preparation

How can I make a model that looks at site fidelity? First I would need to modify the data some more, creating a categorical variable `fidelity` for individuals that have returned to the same pool across years "yes" or have not returned to the same pool across years "no". Then I could try to predict the response of that categorical variable (`fidelity`) based on salinity? (high/low category or just numerical)...

```{r set-up-data-for-model-fidelity}
sals_sight_twice <- sals %>% 
  group_by(pit, year) %>% 
  arrange(date) %>% 
  slice(1) %>% #Get the first row (pool) for each year that the salamander was observed
  ungroup() %>%
  group_by(pit) %>%
  summarise(sightings = n()) %>%
  filter(sightings > 1) %>% #Filter for salamanders observed across two years
  pull(pit) 

sals_fidelity <- sals %>% 
  filter(pit %in% sals_sight_twice) %>%
  group_by(pit, year) %>% 
  slice(1) %>%
  ungroup() %>%
  group_by(pit) %>%
  summarise(pools = toString(pool)) %>%
  separate_wider_delim(cols = pools, delim = ", ", names = c("pool1", "pool2")) %>%
  mutate(fidelity = case_when(pool1 == pool2 ~ "yes",
                               TRUE ~ "no"))

environ22 <- sals %>% 
  filter(pit %in% sals_sight_twice) %>%
  group_by(pit, year) %>% 
  slice(1) %>% 
  filter(year == "2022") %>% 
  rename(pool1_salinity = salinity,
         pool1_temp = pool_temp,
         pool1 = pool,
         year1 = year,
         date1 = date) %>% 
  select(date1, year1, pool1, pit, pool1_salinity, pool1_temp)

environ23 <- sals %>% 
  filter(pit %in% sals_sight_twice) %>%
  group_by(pit, year) %>% 
  slice(1) %>% 
  filter(year == "2023") %>% 
  rename(pool2_salinity = salinity,
         pool2_temp = pool_temp,
         pool2 = pool,
         year2 = year,
         date2 = date) %>% 
  select(date2, year2, pool2, pit, pool2_salinity, pool2_temp)

fidelity <- left_join(sals_fidelity, environ22, by = c("pool1", "pit"))
fidelity <- left_join(fidelity, environ23, by = c("pool2", "pit"))

fidelity <- fidelity %>% 
  mutate(salinity_change = pool2_salinity - pool1_salinity) %>% #Should it be pool2 - pool1 or the other way around?
  mutate(temp_change = pool2_temp - pool1_temp) %>% 
  drop_na(pit) %>% 
  mutate(salinity_magnitude = abs(salinity_change)) %>% 
  mutate(temp_magnitude = abs(temp_change)) %>% 

##Create direction variables for both salinity & temp change where, if the value was negative, direction = decrease and if the value was positive, direction = increase

  mutate(sal_change_direction = if_else(
    salinity_change < 0, "decrease", "increase")) %>% 
  mutate(temp_change_direction = if_else(
    temp_change < 0, "decrease", "increase"))

write_csv(fidelity, "/Users/georgialattig/salamanders by the sea/data/tidy_data/fidelity.csv")
```

My data consist of 96 individual salamanders that were observed across two years. All 96 individuals were PIT-tagged in 2022 and returned to Otter Point in 2023. "Site fidelity" for the sake of this analysis is determined by whether the individual was observed within the same pool on the first night it was observed/found across both seasons. i.e. Did the individual return to the same first pool in 2022 and 2023. With these data, I will model fidelity in response to predictors of change in salinity and change in temperature, as well as see whether the fidelity depends on the pools themselves because there may be other characteristics of the breeding pools that could affect fidelity that aren't reflected in the environmental conditions. 

### Exploratory Data Visualization

```{r fidelity-first-year-barplot, include = FALSE}
fidelity %>% 
  right_join(sals, by = "pit") %>% 
  drop_na(pit, fidelity) %>% 
ggplot(aes(x = fidelity,
                  fill = first_year)) +
  geom_bar() + 
  scale_fill_manual(values = c("yellow2","pink","orange2","darkred","cornflowerblue")) +
  labs(title = "Site Fidelity of Individual Spotted Salamanders",
       subtitle = "Observed Across 2 Seasons",
       fill = "First Year Observed",
       x = "Fidelity",
       y = "Number of Observations") #Fuss with the y axis some more because right now it is including duplicate observations
```

```{r fidelity-density-plot, include = FALSE}
fidelity %>% 
  ggplot(aes(salinity_change, fidelity, fill = fidelity)) + 
  geom_density_ridges(alpha = 0.5) +
  labs(title = "Salinity Changes Experienced by 96 Individually-Tagged Salamanders",
       subtitle = "Faceted by Whether or Not the Individual Exhibited Site Fidelity",
       fill = "Fidelity",
       x = "Change in Salinity",
       y = "Number of Individuals")
```

```{r fidelity-salinity-change-histogram, include = TRUE}
fidelity %>% 
  ggplot(aes(salinity_change))+
  geom_histogram(binwidth = 2) +
  facet_wrap(~ fidelity) +
  labs(title = "Salinity Changes Experienced by 96 Individually-Tagged Salamanders Across 2 Seasons",
       subtitle = "Faceted by Whether or Not the Individual Exhibited Site Fidelity",
       x = "Change in Salinity (ppt)",
       y = "Number of Individuals")
```
```{r fidelity-temp-change-histogram, include = TRUE}
fidelity %>% 
  ggplot(aes(temp_change)) +
  geom_histogram(binwidth = 2) +
  facet_wrap(~ fidelity) +
  labs(title = "Temperature Changes Experienced by 96 Individually-Tagged Salamanders Across 2 Seasons",
       subtitle = "Faceted by Whether or Not the Individual Exhibited Site Fidelity",
       x = "Change in Temperature (C)",
       y = "Number of Individuals")
```

```{r fidelity-salinity-temp-change-scatterplots, include = TRUE}
fidelity %>% 
  group_by(pool1) %>% 
  count()

fidelity %>% 
   mutate(pool1 = factor(pool1, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16"))) %>%
  filter(pool1 %in% c("0","1","3","5","6","7","11","12","13","15")) %>% 
  ggplot(mapping = aes(x = salinity_change,
                     y = temp_change,
                     colour = fidelity)) +
  geom_jitter() +
  facet_wrap(~ pool1) +
  labs(title = "Salinity Change and Temperature Change of Spotted Salamander Breeding Pools",
       subtitle = "Used by Salamanders Returning to the Site Across Two Years",
       x = "Salinity Change (ppt)", y = "Temperature Change (C)",
       colour = "Fidelity") +
  theme_bw() +
  scale_colour_manual(values = c("darkred", "cornflowerblue"))

## Zoom in on values close to zero`
  
fidelity %>% 
   mutate(pool1 = factor(pool1, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16"))) %>%
  filter(pool1 %in% c("0","1","3","5","6","7","11","12","13","15")) %>% 
  ggplot(mapping = aes(x = salinity_change,
                     y = temp_change,
                     colour = fidelity)) +
  geom_jitter() +
  coord_cartesian(xlim = c(-0.5, 0.3), ylim = c(-2, 10)) +
  facet_wrap(~ pool1) +
  labs(title = "Salinity Change and Temperature Change of Spotted Salamander Breeding Pools",
       subtitle = "Used by Salamanders Returning to the Site Across Two Years",
       x = "Salinity Change (ppt)", y = "Temperature Change (C)",
       colour = "Fidelity") +
  theme_bw() +
  scale_colour_manual(values = c("darkred", "cornflowerblue"))
```

```{r salinity-change-boxplot, include = FALSE}
ggplot(fidelity, aes(x = salinity_change)) +
  geom_boxplot() +
  coord_flip() +
  facet_wrap(~ fidelity) +
  labs(title = "Salinity Changes Experienced by 96 Individual A.maculatum",
       subtitle = "Who Returned in 2023",
       x = "Salinity Change")
```

```{r fidelity-pool-barplots, include = TRUE}
fidelity %>% 
  mutate(pool1 = factor(pool1, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16"))) %>%
  mutate(sal = 1) %>% 
  ggplot(aes(x = pool1, y = sal, fill = fidelity)) +
  geom_bar(position="fill", stat="identity") +
  scale_fill_manual(values = c("darkred", "cornflowerblue")) +
  theme_minimal(base_line_size = 7/22) +
  labs(title = "Proportion of Site Fidelity Exhibited by Individual Spotted Salamanders",
       subtitle = "Across 15 Breeding Pools",
       fill = "Fidelity",
       x = "Breeding Pool",
       y = "Number of Salamanders")

fidelity %>% 
  mutate(pool1 = factor(pool1, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16"))) %>%
  mutate(sal = 1) %>% 
  ggplot(aes(x = pool1, y = sal, fill = fidelity)) +
  geom_bar(position="stack", stat="identity") +
  scale_fill_manual(values = c("darkred", "cornflowerblue")) +
  theme_minimal(base_line_size = 7/22) +
  labs(title = "Site Fidelity of Individual Spotted Salamanders",
       subtitle = "Across 15 Breeding Pools",
       fill = "Fidelity",
       x = "Initial Breeding Pool",
       y = "Number of Salamanders")

## Potentially add labels or change tick marks to make the differences in scale more apparent
```
Below I looked into creating a heatmap that would plot pathways between pools (pool 1 on x axis and pool 2 on y axis). Tabling this for now depending on the route I go with the logistic regression modeling. I have a feeling I'm going to focus more on the environmental conditions of pools in relation to fidelity this time around, less on the pools themselves.

```{r pool-fidelity-heatmap, include = FALSE}
fidelity %>% 
  group_by(pool1) %>% 
  count

fidelity %>% 
  group_by(pool2) %>% 
  count

hmfidel <- fidelity %>% 
  group_by(pool1, pool2) %>% 
  count() %>% 
  mutate(connections = n) %>% 
  mutate(connections = as.numeric(connections)) 

hmfidelmat <- as.matrix(hmfidel)

#Need to work harder to get this into proper matrix format for heatmapping

heatmap(hmfidelmat)
```

### Fitting a Model

Some of my target questions with the `fidelity` dataset to explore with models are as follows: 
1. Does the site fidelity of individual salamanders vary with a change in the environmental factors of their breeding pools (salinity, temperature)?
2. Does the site fidelity of individual salamanders vary with the the direction of change in salinity/temperature (increase/decrease) OR the magnitude of change in salinity/temperature?
3. Does the site fidelity of individual salamanders vary with the salinity/temperature of the breeding pool they experienced in that pool during the previous season?
4. Does the site fidelity of individual salamanders vary with the breeding pool they were found in first during the previous season?

```{r model-split-data}
set.seed(2112)
fidel_split <- initial_split(fidelity)

fidel_train <- training(fidel_split)
dim(fidel_train)

fidel_test  <- testing(fidel_split)
dim(fidel_test)
```

Below I set up model workflows for many different but related models which I was interested in running and comparing, to see how well each explained the variation in my data.

```{r build-model}
fidelity_mod <- logistic_reg() %>%
  set_engine("glm") 
```

```{r salinity_change model}
fidelity_rec <- recipe(factor(fidelity) ~ salinity_change + temp_change, 
                       data = fidel_train, family = "binomial")
```

```{r temperature_change model}
fidelity_rec <- recipe(factor(fidelity) ~ salinity_change + temp_change, 
                       data = fidel_train, family = "binomial")
```

```{r salinity_mag_change*direction model}
fidelity_rec <- recipe(factor(fidelity) ~ salinity_change + temp_change, 
                       data = fidel_train, family = "binomial")
```

```{r temp_mag_change_model*direction model}
fidelity_rec <- recipe(factor(fidelity) ~ salinity_change + temp_change, 
                       data = fidel_train, family = "binomial")
```

```{r pool1_salinity*salinity_change model}
fidelity_rec <- recipe(factor(fidelity) ~ salinity_change + temp_change, 
                       data = fidel_train, family = "binomial")
```


```{r pool1_temp*temp_change model}
```






```{r}

fidelity_rec <- recipe(factor(fidelity) ~ salinity_change + temp_change, 
                       data = fidel_train, family = "binomial")
# Do my response variables make sense? Should I standardize them somehow because right now there are both positive and negative differences... How would I interpret the model then? I could create another categorical variable "increase" or "decrease"..?

fidel_wflow <- workflow() %>%
  add_model(fidelity_mod) %>%
  add_recipe(fidelity_rec)

# fidel_fit <- fidel_wflow %>%
#   fit(data = fidel_train)
# tidy(fidel_fit)

# fidelity_main_fit <- linear_reg() %>%
#   set_engine("lm") %>%
#   fit(fidelity ~ salinity + pool_temp + pool, data = sals)
# tidy(fidelity_main_fit)
```

